---
title: "Case 1 Extra Credit"
author: "Daisy Fang"
date: "2/23/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r load-packages}
library(tidyverse)
library(survival)
```

```{r load-data}
dig <- read_csv("dig.csv")
```

Any discrepancy in replication results are noted. 

# Table 1

## Count by treatment
```{r}
# mutate name of treatment
dig <- dig %>% 
  mutate(treatment = case_when(TRTMT == 0 ~ "Placebo",
                         TRUE ~ "Digoxin"))

dig$treatment <- relevel(as.factor(dig$treatment), ref = "Placebo")

# count by treatment
dig %>% 
  count(treatment) %>% 
  kable()
```

## Age
```{r}
dig %>% 
  group_by(treatment) %>% 
  summarize(mean_age = round(mean(AGE), 1),
            sd_age = round(sd(AGE), 1)) %>% 
  kable()
```

## Ejection fraction
```{r}
dig %>% 
  group_by(treatment) %>% 
  summarize(mean_ejf = round(mean(EJF_PER), 1),
            sf_ejf = round(sd(EJF_PER), 1)) %>%  # Digoxin SD is 8.8 instead of 8.9 
  kable()
```

The standard deviation value for the digoxin group is 8.8 instead of 8.9.

## Median duration of CHF
```{r}
dig %>% 
  group_by(treatment) %>% 
  summarize(median_chf = median(CHFDUR, na.rm = TRUE)) %>%  # removed 14 NA's
  kable()
```

## Female Sex
```{r}
dig %>% 
  group_by(treatment, SEX) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

## Nonwhite race
```{r}
dig %>% 
  group_by(treatment, RACE) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

The value for digoxin group is 14.3% instead of 14.4%.

## Age >70 yr
```{r}
dig %>% 
  mutate(age_over_70 = case_when(AGE > 70 ~ "yes",
                                 TRUE ~ "no")) %>% 
  group_by(treatment, age_over_70) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

## Method of assessing ejection fraction
```{r}
dig %>% 
  mutate(method = case_when(EJFMETH == 1 ~ "Radionuclide ventriculography",
                            EJFMETH == 2 ~ "Two-dimensional echocardiography",
                            TRUE ~ "Contrast angiography")) %>% 
  group_by(treatment, method) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(method, percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  kable()
```

## Cardiothoracic ratio >0.55
```{r}
dig %>% 
  mutate(ratio_over_0.55 = case_when(CHESTX > 0.55 ~ "yes",
                                 TRUE ~ "no")) %>% 
  group_by(treatment, ratio_over_0.55) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

## NYHA class
```{r}
dig %>% 
  mutate(class = case_when(FUNCTCLS == 1 ~ "I",
                           FUNCTCLS == 2 ~ "II",
                           FUNCTCLS == 3 ~ "III",
                           FUNCTCLS == 4 ~ "IV")) %>% 
  filter(!is.na(class)) %>% 
  group_by(treatment, class) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(class, percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  kable()
```

## No. of signs or symptoms of CHF
```{r}
dig %>% 
  mutate(number = case_when(NSYM == 0 ~ "0",
                            NSYM == 1 ~ "1",
                           NSYM == 2 ~ "2",
                           NSYM == 3 ~ "3",
                           NSYM == 4 ~ ">=4")) %>% 
  group_by(treatment, number) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(number, percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  kable()
```
The value for the digoxin group over 4 events is 80.2 instead of 80.1.

## Medical history
### Previous myocardial infarction:
```{r}
dig %>% 
  group_by(treatment, PREVMI) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

### Current angina:
```{r}
dig %>% 
  group_by(treatment, ANGINA) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

### Diabetes:
```{r}
dig %>% 
  group_by(treatment, DIABETES) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

### Hypertension:
```{r}
dig %>% 
  group_by(treatment, HYPERTEN) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

## Previous digoxin use
```{r}
dig %>% 
  group_by(treatment, DIGUSE) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```


## Primary cause of CHF
### Ischemic:
```{r}
dig %>% 
  mutate(ischemic = case_when(CHFETIOL == 1 ~ "yes",
                              TRUE ~ "no")) %>% 
  group_by(treatment, ischemic) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```
The placebo value is 70.5 instead of 70.4.

### Nonischemic:
```{r}
dig %>% 
  filter(!is.na(CHFETIOL)) %>% 
  mutate(nonischemic = case_when(CHFETIOL == 1 ~ "no",
                              TRUE ~ "yes")) %>% 
  group_by(treatment, nonischemic) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

#### Nonischemic - Idiopathic:
```{r}
dig %>% 
  filter(!is.na(CHFETIOL)) %>% 
  mutate(idiopathic = case_when(CHFETIOL == 4 ~ "yes",
                              TRUE ~ "no")) %>% 
  group_by(treatment, idiopathic) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

The placebo group value is 14.2 instead of 14.1.

#### Nonischemic - Hypertensive:
```{r}
dig %>% 
  filter(!is.na(CHFETIOL)) %>% 
  mutate(hypertensive = case_when(CHFETIOL == 2 ~ "yes",
                              TRUE ~ "no")) %>% 
  group_by(treatment, hypertensive) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

#### Nonischemic - Other:
```{r}
dig %>% 
  filter(!is.na(CHFETIOL)) %>% 
  mutate(other = case_when(CHFETIOL %in% c(3, 5, 6) ~ "yes",
                              TRUE ~ "no")) %>% 
  group_by(treatment, other) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```
The digoxin group value is 5.5 instead of 5.4.

## Concomitant medications
### Diuretics:
```{r}
dig %>% 
  mutate(diuretics = case_when(DIURETK == 1 | DIURET == 1 ~ "yes",
                               TRUE ~ "no")) %>% 
  group_by(treatment, diuretics) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

### ACE inhibitors:
```{r}
dig %>% 
  group_by(treatment, ACEINHIB) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

### Nitrates:
```{r}
dig %>% 
  group_by(treatment, NITRATES) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

The value for digoxin group is 42.2 instead of 42.1.

### Other vasodilators:
```{r}
dig %>% 
  group_by(treatment, VASOD) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  slice(2, 4) %>% 
  kable()
```

## Daily dose of study medication prescribed 
```{r}
dig %>% 
  mutate(dose = DIGDOSE) %>% 
  group_by(treatment, dose) %>% 
  count() %>% 
  group_by(treatment) %>% 
  summarize(dose, percent_of_patients = round(n/sum(n)*100, 1)) %>% 
  filter(dose %in% c(0.125, 0.250, 0.375, 0.500)) %>% 
  kable()
```

The placebo 0.250mg value is 70.1 instead of 70.0.

# Table 4

## Ejection fraction
```{r}
ef <- dig %>% 
  mutate(ejection_fraction = case_when(EJF_PER < 25 ~ "<0.25",
                                       TRUE ~ "0.25-0.45")) %>% 
  count(treatment, ejection_fraction) %>% 
  rename(total_cnt = n)

dig %>% 
  filter(DWHF == 1) %>% 
  mutate(ejection_fraction = case_when(EJF_PER < 25 ~ "<0.25",
                                       TRUE ~ "0.25-0.45")) %>% 
  count(treatment, ejection_fraction) %>% 
  rename(event_cnt = n) %>% 
  left_join(ef) %>% 
  summarise(treatment, ejection_fraction, event_cnt, total_cnt, percent = round(event_cnt/total_cnt*100, 1)) %>% 
  kable()
```

### Ejection fraction 0.25-0.45
Absolute difference: $32.3-27.0=-5.3$

95% CI:$(-8.0, -2.7)$
```{r}
prop.test(x = c(613,735), n = c(2270, 2273), p = NULL,
          alternative = "two.sided",
          conf.level = 0.95, correct = FALSE)
```

Risk ratio: $0.80$

95% CI: $(0.72, 0.89)$

```{r}
efbt <- dig %>%   
  mutate(ejection_fraction = case_when(EJF_PER < 25 ~ "<0.25",
                                       TRUE ~ "0.25-0.45")) %>% 
  filter(ejection_fraction == "0.25-0.45")

mod <- coxph(Surv(DWHFDAYS, DWHF) ~ treatment, data = efbt)
summary(mod)
```

### Ejection fraction <0.25
Absolute difference: $49.2-38.0=-11.2$

95% CI:$(-15.3, -7.2)$
```{r}
prop.test(x = c(428,556), n = c(1127, 1130), p = NULL,
          alternative = "two.sided",
          conf.level = 0.95, correct = FALSE)
```

Risk ratio: $0.68$

95% CI: $(0.60, 0.77)$

```{r}
efless <- dig %>%   
  mutate(ejection_fraction = case_when(EJF_PER < 25 ~ "<0.25",
                                       TRUE ~ "0.25-0.45")) %>% 
  filter(ejection_fraction != "0.25-0.45")

mod <- coxph(Surv(DWHFDAYS, DWHF) ~ treatment, data = efless)
summary(mod)
```

## Previous use of digoxin

```{r}
pdu <- dig %>% 
  mutate(previous_digoxin_use = case_when(DIGUSE == 0 ~ "No",
                                       TRUE ~ "Yes")) %>% 
  count(treatment, previous_digoxin_use) %>% 
  rename(total_cnt = n)

dig %>% 
  filter(DWHF == 1) %>% 
  mutate(previous_digoxin_use = case_when(DIGUSE == 0 ~ "No",
                                       TRUE ~ "Yes")) %>% 
  count(treatment, previous_digoxin_use) %>% 
  rename(event_cnt = n) %>% 
  left_join(pdu) %>% 
  summarise(treatment, previous_digoxin_use, event_cnt, total_cnt, percent = round(event_cnt/total_cnt*100, 1)) %>% 
  kable()

```

### Existing previous use of digoxin
Absolute difference: $45.3-36.7=-8.6$

95% CI:$(-12.1, -5.1)$
```{r}
prop.test(x = c(550,688), n = c(1498, 1519), p = NULL,
          alternative = "two.sided",
          conf.level = 0.95, correct = FALSE)
```

Risk ratio: $0.74$

95% CI: $(0.66, 0.83)$

```{r}
yesuse <- dig %>%   
  mutate(previous_digoxin_use = case_when(DIGUSE == 0 ~ "No",
                                       TRUE ~ "Yes")) %>% 
  filter(previous_digoxin_use == "Yes")

mod <- coxph(Surv(DWHFDAYS, DWHF) ~ treatment, data = yesuse)
summary(mod)
```

### No previous use of digoxin
Absolute difference: $32.0-25.9=-6.2$

95% CI:$(-9.0, -3.3)$
```{r}
prop.test(x = c(491,603), n = c(1899, 1884), p = NULL,
          alternative = "two.sided",
          conf.level = 0.95, correct = FALSE)
```

Risk ratio: $0.77$

95% CI: $(0.68, 0.86)$

```{r}
nouse <- dig %>%   
  mutate(previous_digoxin_use = case_when(DIGUSE == 0 ~ "No",
                                       TRUE ~ "Yes")) %>% 
  filter(previous_digoxin_use == "No")

mod <- coxph(Surv(DWHFDAYS, DWHF) ~ treatment, data = nouse)
summary(mod)
```

## Cause of heart failure
```{r}
chf <- dig %>% 
  filter(!is.na(CHFETIOL)) %>% 
  mutate(ischemic = case_when(CHFETIOL == 1 ~ "Yes",
                              TRUE ~ "No")) %>% 
  count(treatment, ischemic) %>% 
  rename(total_cnt = n)

dig %>% 
  filter(DWHF == 1) %>% 
  filter(!is.na(CHFETIOL)) %>% 
  mutate(ischemic = case_when(CHFETIOL == 1 ~ "Yes",
                              TRUE ~ "No")) %>% 
  count(treatment, ischemic) %>% 
  rename(event_cnt = n) %>% 
  left_join(chf) %>% 
  summarise(treatment, ischemic, event_cnt, total_cnt, percent = round(event_cnt/total_cnt*100, 1)) %>% 
  kable()
```

### Ischemic cause of heart failure
Absolute difference: $36.4-30.4=-6.0$

95% CI:$(-8.7, -3.3)$
```{r}
prop.test(x = c(731, 873), n = c(2405, 2398), p = NULL,
          alternative = "two.sided",
          conf.level = 0.95, correct = FALSE)
```

Risk ratio: $0.79$

95% CI: $(0.72, 0.88)$

```{r}
ischemic <- dig %>%   
  filter(!is.na(CHFETIOL)) %>% 
  mutate(ischemic = case_when(CHFETIOL == 1 ~ "Yes",
                              TRUE ~ "No")) %>% 
  filter(ischemic == "Yes")

mod <- coxph(Surv(DWHFDAYS, DWHF) ~ treatment, data = ischemic)
summary(mod)
```

### Nonischemic cause of heart failure
Absolute difference: $41.5-31.1=-10.3$

95% CI:$(-14.5, -6.1)$
```{r}
prop.test(x = c(306, 413), n = c(983, 996), p = NULL,
          alternative = "two.sided",
          conf.level = 0.95, correct = FALSE)
```

Risk ratio: $0.67$

95% CI: $(0.58, 0.77)$

```{r}
nonischemic <- dig %>%   
  filter(!is.na(CHFETIOL)) %>% 
  mutate(ischemic = case_when(CHFETIOL == 1 ~ "Yes",
                              TRUE ~ "No")) %>% 
  filter(ischemic == "No")

mod <- coxph(Surv(DWHFDAYS, DWHF) ~ treatment, data = nonischemic)
summary(mod)
```


## Cardiothoracic ratio
```{r}
cr <- dig %>% 
  mutate(cardiothoracic_ratio = case_when(CHESTX > 0.55 ~ "Yes",
                                 TRUE ~ "No")) %>% 
  count(treatment, cardiothoracic_ratio) %>% 
  rename(total_cnt = n)

dig %>% 
  filter(DWHF == 1) %>% 
  mutate(cardiothoracic_ratio = case_when(CHESTX > 0.55 ~ "Yes",
                                 TRUE ~ "No")) %>% 
  count(treatment, cardiothoracic_ratio) %>% 
  rename(event_cnt = n) %>% 
  left_join(cr) %>% 
  summarise(treatment, cardiothoracic_ratio, event_cnt, total_cnt, percent = round(event_cnt/total_cnt*100, 1)) %>% 
  kable()
```

The count for cardiothoracic ratio <=0.55 in the digoxin group is 2221 instead of 2220, though the proportions are not different from the paper after rounding.

### Cardiothoracic ratio <= 0.55
Absolute difference: $32.4-27.0=-5.4$

95% CI:$(-8.1, -2.7)$
```{r}
prop.test(x = c(600, 724), n = c(2221, 2233), p = NULL,
          alternative = "two.sided",
          conf.level = 0.95, correct = FALSE)
```

Risk ratio: $0.79$

95% CI: $(0.71, 0.88)$

```{r}
cr_less <- dig %>%   
  mutate(cardiothoracic_ratio = case_when(CHESTX > 0.55 ~ "Yes",
                                 TRUE ~ "No")) %>% 
  filter(cardiothoracic_ratio == "No")

mod <- coxph(Surv(DWHFDAYS, DWHF) ~ treatment, data = cr_less)
summary(mod)
```

### Cardiothoracic ratio > 0.55
Absolute difference: $48.5-37.5=-11$

95% CI:$(-14.9, -7.0)$
```{r}
prop.test(x = c(441, 567), n = c(1176, 1170), p = NULL,
          alternative = "two.sided",
          conf.level = 0.95, correct = FALSE)
```

Risk ratio: $0.69$

95% CI: $(0.61, 0.78)$

```{r}
cr_more <- dig %>%   
  mutate(cardiothoracic_ratio = case_when(CHESTX > 0.55 ~ "Yes",
                                 TRUE ~ "No")) %>% 
  filter(cardiothoracic_ratio == "Yes")

mod <- coxph(Surv(DWHFDAYS, DWHF) ~ treatment, data = cr_more)
summary(mod)
```


## NYHA class
```{r}
nyha <- dig %>% 
  filter(!is.na(FUNCTCLS)) %>% 
  mutate(class = case_when(FUNCTCLS %in% c(1, 2) ~ "I or II",
                           FUNCTCLS %in% c(3, 4) ~ "III or IV")) %>% 
  count(treatment, class) %>% 
  rename(total_cnt = n)

dig %>% 
  filter(DWHF == 1) %>% 
  filter(!is.na(FUNCTCLS)) %>% 
  mutate(class = case_when(FUNCTCLS %in% c(1, 2) ~ "I or II",
                           FUNCTCLS %in% c(3, 4) ~ "III or IV")) %>% 
  count(treatment, class) %>% 
  rename(event_cnt = n) %>% 
  left_join(nyha) %>% 
  summarise(treatment, class, event_cnt, total_cnt, percent = round(event_cnt/total_cnt*100, 1)) %>% 
  kable()
```

### NYHA class I or II
Absolute difference: $32.2-26.4=-5.8$

95% CI:$(-8.4, -3.1)$
```{r}
prop.test(x = c(601, 739), n = c(2275, 2296), p = NULL,
          alternative = "two.sided",
          conf.level = 0.95, correct = FALSE)
```

Risk ratio: $0.78$

95% CI: $(0.70, 0.87)$

```{r}
nyha_12 <- dig %>%   
  filter(!is.na(FUNCTCLS)) %>% 
  mutate(class = case_when(FUNCTCLS %in% c(1, 2) ~ "I or II",
                           FUNCTCLS %in% c(3, 4) ~ "III or IV")) %>% 
  filter(class != "III or IV")

mod <- coxph(Surv(DWHFDAYS, DWHF) ~ treatment, data = nyha_12)
summary(mod)
```

### NYHA class III or IV
Absolute difference: $50.0-39.2=-10.8$

95% CI:$(-14.9, -6.7)$, this is slightly different from the $(-14.9, -6.7)$ as reported
```{r}
prop.test(x = c(438, 552), n = c(1118, 1105), p = NULL,
          alternative = "two.sided",
          conf.level = 0.95, correct = FALSE)
```

Risk ratio: $0.70$

95% CI: $(0.61, 0.79)$

```{r}
nyha_34 <- dig %>%   
  filter(!is.na(FUNCTCLS)) %>% 
  mutate(class = case_when(FUNCTCLS %in% c(1, 2) ~ "I or II",
                           FUNCTCLS %in% c(3, 4) ~ "III or IV")) %>% 
  filter(class == "III or IV")

mod <- coxph(Surv(DWHFDAYS, DWHF) ~ treatment, data = nyha_34)
summary(mod)
```

## Overall study population
```{r}
tot <- dig %>% 
  count(treatment) %>% 
  rename(total_cnt = n)

dig %>% 
  filter(DWHF == 1) %>% 
  count(treatment) %>% 
  rename(event_cnt = n) %>% 
  left_join(tot) %>% 
  summarise(treatment, event_cnt, total_cnt, percent = round(event_cnt/total_cnt*100, 1)) %>% 
  kable()
```

Absolute difference: $37.9-30.6=-7.3$

95% CI:$(-9.5, -5.0)$
```{r}
prop.test(x = c(1041, 1291), n = c(3397, 3403), p = NULL,
          alternative = "two.sided",
          conf.level = 0.95, correct = FALSE)
```

Risk ratio: $0.75$

95% CI: $(0.69, 0.82)$

```{r}
mod <- coxph(Surv(DWHFDAYS, DWHF) ~ treatment, data = dig)
summary(mod)
```



